---
title: "HW5_Do_Quyen"
author: "Quyen Do"
date: "September 21, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 3
My thoughts on what makes a good figure
It should give you some new information or insights about the data at hand

## Problem 4
a. A function computing the proportion of successes in a vector

```{r count_success}
## TODO: run microbenchmark on different count_success function

count_success <- function (vect, value = 1) {
  # Compute the proportion of successes in a vector
  
  # Args:
  # vect: the vector on which the proportion of successes will be computed
  # value: the value presented "success" value in the vector. Default value is 1
  
  #Return:
  # A real number from 0 to 1
  
  length(vect[which(vect==value)])/length(vect)
}
```
b. Create a simuluated matrix
```{r binom_matrix}
set.seed(12345)
P4b_data <- matrix(rbinom(10,1,prob =(30:40)/100),nrow = 10, ncol =10)
```
c. Checking the proportion of success
```{r}
# Calculate the proportion of success across matrix row
prop_row <- apply(P4b_data,1,count_success)
prop_col <- apply(P4b_data,2,count_success)

prop_mat <- matrix(c(prop_row,prop_col),nrow=2,ncol=10,byrow = TRUE, dimnames = list(c("By Row","By Col"),1:10))
prop_mat
```
The matrix of the simulated binomial using the code in b didn't produce the data as the intentionn. Instead of applying different probabilities to draw success among 10 rows of the matrix, the function seems to apply $p=1$ to all the rows instead.

d.
```{r}
simulate_binom <- function(probability) {
  # Simulate 10 random binomial variables 
  #  of n = 10 and given probability
  
  # Args:
  # probability: the probability for the binomial distribution
  
  #Return:
  # a vector containing 10 RVs drawn from binomial distribution
  
  print(probability)
  return(rbinom(10, 1, prob = probability))
}

# A vector of probability
prob_vect <- (30:40)/100

# apply simulate_binom on each element of prob_vect
correct_mat <- sapply(prob_vect,simulate_binom)

# Calculate the proportion of success
# across rows and columns of correct_mat
prop_row2 <- apply(correct_mat,1,count_success)
prop_col2 <- apply(correct_mat,2,count_success)

prop_mat2 <- matrix(c(prop_row2,prop_col2),nrow=2,ncol=10,byrow = TRUE, dimnames = list(c("By Row","By Col"),1:10))
prop_mat2
```


